### 2019-07-06 Memoria/Chapters/ExperimentalSetup.tex
Cambios en memoria y en paper DILS

In this chapter we present the experimental setup used to compare the two new proposals introduced in Chapter \ref{ch:NewProposals} with the current state of the art. We will start by briefly introducing the state-of-the-art methods in Section \ref{sec:SOTA}. Afterwards we specify the dataset used in experimentation (Section \ref{sec:Datasets}), as well as the constraint sets (Section \ref{sec:ConstGen}) and the results evaluation method (Section \ref{sec:EvalMet}). In Section \ref{sec:ValidtnMethod} we introduce the validation method we will apply to the results. We also include the required specifications to replicate the experiments in Section \ref{sec:Calibration}.

@@

The first adaptation of a classic clustering method for constrained clustering was proposed in \cite{wagstaff2001constrained}. It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL. This method was named \acf{COPKM}, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.

In \cite{antoine2012cecm} \acf{CECM}, a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods. The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function. This method redefines constraints from the point of view of belief functions and includes them in the cost function.

A modification of the Constrained Vector Quantization Error algorithm (CVQE \cite{davidson2005clustering}) is proposed in \cite{pelleg2007k}. The CVQE algorithm proved to produce high quality results, at the cost of a very high computational complexity. \acf{LCVQE} introduces a modification of the cost function of CVQE to make it more intuitive and less computationally complex. The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraint sets.

@@

For our experiments we will compare the results obtained by \acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$ with those obtained by state-of-the-art methods and \acs{BRKGA}$_{CC}$ over 20 datasets and 3 constraint sets for each one of them. Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}. We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}. Table \ref{tab:datasets} displays a summary of every dataset.

@@

Since we have the true labels associated with each dataset, we will use the method proposed in \cite{wagstaff2001constrained} to generate artificial constraint sets. This method consists of randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.

@@

\begin{table}[!h]
	\centering
	%\setlength{\tabcolsep}{7pt}
	%\renewcommand{\arraystretch}{1}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lcc c cc c cc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{2}{c}{$CS_{10}$} && \multicolumn{2}{c}{$CS_{15}$} && \multicolumn{2}{c}{$CS_{20}$} \\
			\cline{2-3} \cline{5-6} \cline{8-9}
			& ML & CL && ML & CL && ML & CL \\
			\hline
			Appendicitis & 39 & 16 && 71 & 49 && 164 & 67 \\
			Breast Cancer & 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
			Bupa & 323 & 272 && 699 & 627 && 1201 & 1145 \\
			Circles & 208 & 227 && 502 & 488 && 853 & 917 \\
			Ecoli & 163 & 398 && 357 & 918 && 609 & 1669 \\
			Glass & 52 & 179 && 139 & 389 && 259 & 644 \\
			Haberman & 304 & 161 && 634 & 401 && 1135 & 756 \\
			Hayesroth & 39 & 81 && 102 & 174 && 177 & 319 \\
			Ionosphere & 330 & 300 && 732 & 646 && 1299 & 1186 \\
			Iris & 26 & 79 && 82 & 171 && 136 & 299 \\
			Monk2 & 473 & 473 && 979 & 1101 && 1917 & 1824 \\
			Moons & 200 & 235 && 494 & 496 && 900 & 870 \\

@@

The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it. Note that the greater the number of classes present in the dataset, the fewer ML constraints obtained with the method proposed in \cite{wagstaff2001constrained}. This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases. Figure \ref{fig:IrisConst} displays a visual representation of the constraint sets generated for the Iris dataset.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[$CS_{10}$ \acs{ML} constraints]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML10.png}} 
	\subfloat[$CS_{10}$ \acs{CL} constraints]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL10.png}} \quad
	\subfloat[$CS_{15}$ \acs{ML} constraints]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML15.png}} 
	\subfloat[$CS_{15}$ \acs{CL} constraints]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL15.png}} \quad
	\subfloat[$CS_{20}$ \acs{ML} constraints]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML20.png}} 
	\subfloat[$CS_{20}$ \acs{CL} constraints]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL20.png}} 
	\caption{Visual representation of the three constraints sets the Iris dataset.}
	\label{fig:IrisConst}
\end{figure}

@@

Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this paper. To accomplish this task we just set one of the two partitions given to compute \acs{ARI} as the ground truth labels.

\section{Validation of results} \label{sec:ValidtnMethod}

In order to validate the results which will be presented in Chapter \ref{ch:ExperimentalResults} we will use Bayesian statistical tests, instead of the classic \acf{NHST}. In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of \acs{NHST}, and a new model is proposed for carrying out comparisons researchers are interested in. \textit{"In a nutshell: \acs{NHST} do not answer the question we ask"}. To put it clear, the disadvantages of the \acs{NHST} that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?

To start with, \acs{NHST} do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different? Another pitfall of \acs{NHST} is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis. This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.

@@

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	%\resizebox{\textwidth}{!}{
		\begin{tabular}{>{\centering\arraybackslash}c >{\arraybackslash}L{5cm} cc}
			\hline
			Parameter & Meaning & \acs{BRKGA}$_{CC}$ & \acs{SHADE}$_{CC}$ \\
			\hline
			$|P|$ & Population size & 100 & 100 \\
			Evals & Fitness function evaluations & 300000 & 300000 \\
			$P_e$ & Size of the elite set in population & $0.2 * |P|$ & $0.25 * |P|$ \\
			$P_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * |P|$ & - \\
			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets})} \\
			\hline

@@

	\caption[Parameters setup used for BRKGA$_{CC}$ and SHADE$_{CC}$]{Parameters setup used for \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$.}
	\label{tab:paramsSHADE}
\end{table}

@@

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{c L{6cm} c}
			\hline
			Parameter & Meaning & Value \\
			\hline
			Evals & Fitness function evaluations & 300000 \\
			$p_s$ & Segment size for the mutation operator & $0.3 \times n$ \\
			$p_m$ & Maximum number of neighbors generated in each call to the LS procedure & Evals $\times 0.01$ \\
			$p_r$ & Probability that a feature is selected from $m_b$ for recombination & $0.3$ \\
			$\xi$ & Reinitialization method tolerance & $0.2$ \\
			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}) \\
			\hline

@@

	\caption[Parameters setup used for DILS$_{CC}$]{Parameters setup used for \acs{DILS}$_{CC}$.}
	\label{tab:paramsDILS}
\end{table}

@@

	\caption{Parameters setup used for the state-of-the-art algorithms.}
	\label{tab:paramsSOTA}
\end{table}

Parameter values have been assigned following the guidelines of the original creators of the different proposals. Since the evaluation in the experimental stage makes use of a high number of datasets, tuning each parameter specifically for each dataset is not feasible. Indeed, our goal is not optimization in a case-by-case basis but instead a comparison in the most general scenario possible. Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric. All parameters described above are meant to be used for the Python implementation that can be found at \href{https://github.com/GermangUgr}{GitHub}\footnote{https://github.com/GermangUgr}, which will be used to obtain our results.

>>>>

In this chapter we present the experimental setup used to compare the two new proposals introduced in Chapter \ref{ch:NewProposals} with the current state-of-the-art. We start by briefly introducing the state-of-the-art methods in Section \ref{sec:SOTA}. Afterwards, we specify the dataset used in experimentation (Section \ref{sec:Datasets}), as well as the constraint sets (Section \ref{sec:ConstGen}) and the result evaluation method (Section \ref{sec:EvalMet}). The validation method that will be applied to the results in later chapters (Chapter \ref{ch:AnalysisResults} is discussed in Section \ref{sec:ValidtnMethod}). We also include the required configuration to replicate the experiments in Section \ref{sec:Calibration}.

@@

The first adaptation of a classic clustering method for constrained clustering was proposed in \cite{wagstaff2001constrained}. It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known \acs{ML} and \acs{CL}. This method was named \acf{COPKM}, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.

In \cite{antoine2012cecm} \acf{CECM}, a variant of the Evidential c-means (\acsfont{ECM} \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods. The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function. This method redefines constraints from the point of view of belief functions and includes them in the cost function.

A modification of the Constrained Vector Quantization Error algorithm (\acsfont{CVQE} \cite{davidson2005clustering}) is proposed in \cite{pelleg2007k}. The \acsfont{CVQE} algorithm proved to produce high quality results, at the cost of a very high computational complexity. \acf{LCVQE} introduces a modification of the cost function of \acsfont{CVQE} to make it more intuitive and less computationally complex. The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraint sets.

@@

For our experiments we will compare the results obtained by \acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$ with those obtained by state-of-the-art methods and \acs{BRKGA}$_{CC}$ over twenty datasets and three constraint sets for each one of them. Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}. We also include three artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}. Table \ref{tab:datasets} displays a summary of every dataset.

@@

Since we have the true labels associated with each dataset, we will use the method proposed in \cite{wagstaff2001constrained} to generate artificial constraint sets. This method consists of randomly selecting two instances of a dataset, then comparing its labels, and finally setting an \acs{ML} or \acs{CL} constraint depending on whether the labels are the same or different.

@@

\begin{table}[!h]
	\centering
	%\setlength{\tabcolsep}{7pt}
	%\renewcommand{\arraystretch}{1}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lcc c cc c cc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{2}{c}{$CS_{10}$} && \multicolumn{2}{c}{$CS_{15}$} && \multicolumn{2}{c}{$CS_{20}$} \\
			\cline{2-3} \cline{5-6} \cline{8-9}
			& \acs{ML} & \acs{CL} && \acs{ML} & \acs{CL} && \acs{ML} & \acs{CL} \\
			\hline
			Appendicitis & 39 & 16 && 71 & 49 && 164 & 67 \\
			Breast Cancer & 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
			Bupa & 323 & 272 && 699 & 627 && 1201 & 1145 \\
			Circles & 208 & 227 && 502 & 488 && 853 & 917 \\
			Ecoli & 163 & 398 && 357 & 918 && 609 & 1669 \\
			Glass & 52 & 179 && 139 & 389 && 259 & 644 \\
			Haberman & 304 & 161 && 634 & 401 && 1135 & 756 \\
			Hayesroth & 39 & 81 && 102 & 174 && 177 & 319 \\
			Ionosphere & 330 & 300 && 732 & 646 && 1299 & 1186 \\
			Iris & 26 & 79 && 82 & 171 && 136 & 299 \\
			Monk2 & 473 & 473 && 979 & 1101 && 1917 & 1824 \\
			Moons & 200 & 235 && 494 & 496 && 900 & 870 \\

@@

The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it. Note that the greater the number of classes present in the dataset, the fewer \acs{ML} constraints obtained with the method proposed in \cite{wagstaff2001constrained}. This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases. Figure \ref{fig:IrisConst} displays a visual representation of the constraint sets generated for the Iris dataset.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[$CS_{10}$ \acs{ML} constraints.]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML10.png}} 
	\subfloat[$CS_{10}$ \acs{CL} constraints.]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL10.png}} \quad
	\subfloat[$CS_{15}$ \acs{ML} constraints.]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML15.png}} 
	\subfloat[$CS_{15}$ \acs{CL} constraints.]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL15.png}} \quad
	\subfloat[$CS_{20}$ \acs{ML} constraints.]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML20.png}} 
	\subfloat[$CS_{20}$ \acs{CL} constraints.]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL20.png}} 
	\caption{Visual representation of the three constraint sets over the Iris dataset.}
	\label{fig:IrisConst}
\end{figure}

@@

Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this work. To accomplish this task we just set one of the two partitions given to compute \acs{ARI} as the ground truth labels.

\section{Validation of Results} \label{sec:ValidtnMethod}

In order to validate the results which will be presented in Chapter \ref{ch:ExperimentalResults} we will use Bayesian statistical tests, instead of the classic \acf{NHST}. In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of \acs{NHST}, and a new model is proposed for carrying out comparisons researchers are interested in. \textit{"In a nutshell: \acs{NHST} do not answer the question we ask"}. To put it clear, the disadvantages of the \acs{NHST} that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: To reject, or not to reject?

To start with, \acs{NHST} do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: What is the probability that two methods are different? Another pitfall of \acs{NHST} is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis. This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.

@@

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	%\resizebox{\textwidth}{!}{
		\begin{tabular}{c l cc}
			\hline
			Parameter & Meaning & \acs{BRKGA}$_{CC}$ & \acs{SHADE}$_{CC}$ \\
			\hline
			$|P|$ & Population size & 100 & 100 \\
			Evals & Fitness function evaluations & 300000 & 300000 \\
			$P_e$ & \makecell[l]{Size of the elite set in \\ population} & $0.2 * |P|$ & $0.25 * |P|$ \\
			$P_m$ & \makecell[l]{Number of mutants to be \\ introduced in the population in \\ each generation} & $0.2 * |P|$ & - \\
			$p_\text{inherit}$ & \makecell[l]{Probability that a feature is \\ inherited  from an elite parent} & $50\%$ & - \\
			$k$ & \makecell[l]{Output partition number of \\ clusters} & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets})} \\
			\hline

@@

	\caption{Parameter setup used for \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$.}
	\label{tab:paramsSHADE}
\end{table}

@@

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{c l c}
			\hline
			Parameter & Meaning & Value \\
			\hline
			Evals & Fitness function evaluations & 300000 \\
			$p_s$ & \makecell[l]{Segment size for the mutation \\ operator} & $0.3 \times n$ \\
			$p_m$ & \makecell[l]{Maximum number of neighbors \\ generated in each call to the \acs{LS} \\ procedure} & Evals $\times 0.01$ \\
			$p_r$ & \makecell[l]{Probability that a feature is selected \\ from $m_b$ for recombination} & $0.3$ \\
			$\xi$ & Reinitialization method tolerance & $0.2$ \\
			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}) \\
			\hline

@@

	\caption{Parameter setup used for \acs{DILS}$_{CC}$.}
	\label{tab:paramsDILS}
\end{table}

@@

	\caption{Parameter setup used for the state-of-the-art algorithms.}
	\label{tab:paramsSOTA}
\end{table}

Parameter values have been assigned following the guidelines of the original creators of the different proposals. Since the evaluation in the experimental stage makes use of a high number of datasets, tuning each parameter specifically for each dataset is not feasible. Indeed, our goal is not optimization in a case-by-case basis but instead a comparison in the most general scenario possible. Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric. Equivalent results can be obtained with the parameters described above when used in the Python implementation found at \href{https://github.com/GermangUgr}{GitHub}\footnote{https://github.com/GermangUgr}.

### 2019-07-06 Memoria/Chapters/Introduction.tex
Cambios en memoria y en paper DILS

\begin{quotation}{\slshape
		An intelligent being cannot treat every object it sees as unique entity unlike anything else in the universe. It has to put objects in categories so that it may apply its hard-won knowledge about similar objects encountered in the past, to he object at hand.}
		\begin{flushright}
			\textbf{Steven Pinker, How the Mind Works, 1997} 
		\end{flushright}
\end{quotation}

One of the most basic and primitive abilities that human beings are endowed with is being able to group similar objects to produce a classification that is useful to them. For example, they should have been able to realize which objects were edible, which were poisonous, and which would try to kill them.

@@

Just as classification is a basic skill for people in their daily lives, it is also essential in most branches of science. In biology, for example, the classification of different types of organisms has been the subject of study since the beginning of its existence. Aristotle built an elaborated animal classification system, dividing all creatures into two groups: red-blooded and not red-blooded. Later he proposed a subdivision that classified them according to the way in which new individuals came into the world, either alive, by means of eggs, chrysalis, etc.

Following Aristotle, Theophrastus wrote the first document compiling guidelines for the classification of plants. The resulting works were so extensive and detailed that they laid the foundation for research in biology over the following centuries. This work was not replaced until 1737, when Carl von Linné published his work \textit{Genera Plantarum}, from which we extract the following fragment:

@@

Classification of animals and plants has been of major importance in fields such as biology and zoology. In particular, this classification laid the foundation for the development of Darwin's theory of evolution. But it has also been of high relevance in knowledge areas such as chemistry and physics, with the classification of the elements in the periodic table, proposed by Mendeleyev in the 1860s; or in astronomy, with the classification of dwarf or giant stars using the Hertzsprung-Russell guidelines, shown in Figure \ref{fig:intro_HRdiagram}.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.35]{gfx/Intro/HR_diagram} 
	\caption[Hertzsprung-Russell stars classification diagram.]{Hertzsprung-Russell stars classification diagram.}\label{fig:intro_HRdiagram}
\end{figure}

@@

We humans, like many other beings, are capable of learning a multitude of concepts throughout our lives, although generally the learning process is extremely slow. It takes us about six years to be able to begin dedicated and specific learning in primary education, and another eighteen years---generally speaking---to become computer scientists or experts in any other branch of knowledge.

Thus, the human learning process is a slow and tedious process. So, do we want to teach machines to learn the way humans do? Machines have proven to be outstanding at repetitive tasks, as well as they reach calculation speeds that humans can only dream of. It is worth asking if it is possible to develop a learning method for machines that is orders of magnitude faster than the human and equal or more effective. Although there is a major impediment to the development of such a method: we do not know if it exists! The search for a method to avoid the disadvantages of the human learning process is what drives research in the science we know today as \textit{machine learning}.

One of the major differences between human learning and machine learning, and we must admit that to this respect machines again lead us, is the way in which knowledge is transferred. In machines it is enough to copy and paste, that is to say, once a program performs correctly the task for which it has been designed in a computer, it performs it in all of them. Just copy and paste. In humans this is not the case, we cannot extract knowledge from one brain and insert it into another, we don't even know how it is encoded! We humans must use our inefficient input-output devices to transmit knowledge. The following quote, which well reflects the reality of automatic learning, is more than thirty years old, yet still valid:

@@

So far, references have been made to learning without formally defining it. The concept of learning is deeply anchored in human beings. However, when trying to apply learning to computers, which are still no more than glorified calculators, we must give some kind of definition: \textit{Learning denotes changes in the system that are adaptive in the sense that they enable the system to do the same task or tasks drawn from the same population more efficiently and more effectively the next time} \cite{Michalski:2013:MLA:2588013}.

Machine learning is a field derived from computer science, and a branch of \acf{AI}, whose ultimate goal is to develop methods that make it possible for machines to learn. In other words, the techniques embedded in machine learning aim to develop programs that guide computers to learn from a set of examples. Machine learning is part of \acs{AI} in the sense that \acs{AI} is often directed at simulating human beings behavior through computers, so that we can better understand how humans work and help them.

@@

There are several approaches within the field of machine learning, these are classified according to the information that is provided to the machine for it to learn. Thus, following the classification proposed in \cite{ayodele2010types}, we find four different types of learning:

@@

	\item \textbf{Supervised Learning}: complete information is provided for learning. The machine is provided with a set of examples and the class (label) associated with each of them. The goal is to generate a function that approximates the behavior of the function that maps the examples in their respective classes.

	\item \textbf{Unsupervised Learning}: in this case the class to which the examples belong to is not available. The goal is to extract the class from the information contained in the set of examples.

	\item \textbf{Semi-supervised Learning}: it arises from adding incomplete information to unsupervised learning. This information is usually given in the form of a subset of labeled data, but there is no restrictions to this concern.

	\item \textbf{Reinforcement Learning}: the algorithm learns how to act under the influence of its environment, which provides feedback on the actions taken by the algorithm.

@@

Among the four main categories of machine learning methods we find the \acf{SSL}, which is the focus of this work. As we have already mentioned, it is a machine learning paradigm that arises from adding incomplete information to unsupervised learning. We can divide \acs{SSL} methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}. The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances. The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this work \cite{triguero2015self}.

@@

>>>>

\begin{quotation}{\slshape
		An intelligent being cannot treat every object it sees as a unique entity unlike anything else in the universe. It has to put objects in categories so that it may apply its hard-won knowledge about similar objects encountered in the past, to the object at hand.}
		\begin{flushright}
			\textbf{Steven Pinker, How the Mind Works, 1997} 
		\end{flushright}
\end{quotation}

One of the most basic and primitive abilities that human beings are endowed with is being able to group similar objects to produce a classification that is useful to them. For example, they were able to realize which objects were edible, which were poisonous, and which would try to kill them.

@@

Just as classification is a basic skill for people in their daily lives, it is also essential in most branches of science. In biology, for example, the classification of different types of organisms has been the subject of study since the beginning of its existence. Aristotle built an elaborated animal classification system, dividing all creatures into two groups: red-blooded and not red-blooded. Later he proposed a subdivision that classified them according to the way in which new individuals came into the world (either alive, by means of eggs, chrysalis, etc).

Following Aristotle, Theophrastus wrote the first document that compiled guidelines for the classification of plants. The resulting works were so extensive and detailed that they laid the foundation for research in biology over the following centuries. This work was not replaced until 1737, when Carl von Linné published his work \textit{Genera Plantarum}, from which we extract the following fragment:

@@

Classification of animals and plants has been of major importance in fields such as biology and zoology. In particular, this classification laid the foundation for the development of Darwin's theory of evolution. But it has also been highly relevant in knowledge areas such as chemistry and physics, with the classification of the elements in the periodic table, proposed by Mendeleyev in the 1860s; or in astronomy, with the classification of dwarf or giant stars using the Hertzsprung-Russell guidelines, shown in Figure \ref{fig:intro_HRdiagram}.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.35]{gfx/Intro/HR_diagram} 
	\caption{Hertzsprung-Russell stars classification diagram.}\label{fig:intro_HRdiagram}
\end{figure}

@@

We humans, like many other beings, are capable of learning a multitude of concepts throughout our lives, although generally the learning process is extremely slow. It takes us about six years to be able to begin dedicated and specific learning in primary education, and another eighteen years (generally speaking) to become computer scientists or experts in any other branch of knowledge.

Thus, the human learning process is a slow and tedious process. So, do we want to teach machines to learn the way humans do? Machines have proven to be outstanding at repetitive tasks, and they also reach calculation speeds that humans can only dream of. It is worth asking if it is possible to develop a learning method for machines that is orders of magnitude faster than the human method and equally or more effective. Nevertheless, there is a major impediment to the development of such a method: we do not know whether it exists! The search for a method to avoid the disadvantages of the human learning process is what drives research in the field of science we know today as \textit{machine learning}.

One of the major differences between human learning and machine learning, and we must admit that in this respect machines are again ahead of us, is the way in which knowledge is transferred. In machines it is enough to copy and paste: once a program correctly performs the task for which it has been designed in a computer, it can perform it in all of them. Just copy and paste. In humans this is not the case. We cannot extract knowledge from one brain and insert it into another. We don't even know how it is encoded! We humans must use our inefficient input-output devices to transmit knowledge. The following quote, which well reflects the reality of automatic learning, is more than thirty years old, yet still valid:

@@

So far, references have been made to learning without formally providing a definition. The concept of learning is deeply anchored in human beings. However, when trying to apply learning to computers, which are still no more than glorified calculators, the following definition should suffice: \textit{Learning denotes changes in the system that are adaptive in the sense that they enable the system to do the same task or tasks drawn from the same population more efficiently and more effectively the next time} \cite{Michalski:2013:MLA:2588013}.

Machine learning is a field derived from computer science, and a branch of \acf{AI}, whose ultimate goal is to develop methods that make it possible for machines to learn. In other words, the techniques embedded in machine learning aim to develop programs that guide computers in learning from a set of examples. Machine learning is a part of \acs{AI} in the sense that \acs{AI} is often directed at simulating human beings behavior through computers, so that we can better understand how humans work.

@@

There are several approaches within the field of machine learning, which are classified according to the information that is provided to the machine for its learning process. Thus, following the classification proposed in \cite{ayodele2010types}, we have four different types of learning:

@@

	\item \textbf{Supervised Learning}: complete information is available for learning. The machine is provided with a set of examples and the class label associated with each of them. The goal is approximate the behavior of the function that maps the examples to their respective classes.

	\item \textbf{Unsupervised Learning}: in this case the class which the examples belong to is not available. The goal is to deduce the class from the information contained in the set of examples.

	\item \textbf{Semi-supervised Learning}: it arises from adding incomplete information to unsupervised learning. This information is usually given in the form of a subset of labeled data, but there are no restrictions in this regard.

	\item \textbf{Reinforcement Learning}: the algorithm learns how to act within its environment, which provides feedback on the actions taken by the algorithm.

@@

Among the four main categories of machine learning methods we find the \acf{SSL}, which is the focus of this work. As we have already mentioned, it is a machine learning paradigm that arises from adding incomplete information to unsupervised learning. We can divide \acs{SSL} methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}. The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances. The second tries to obtain better-defined clusters incorporating background information to the clustering process; this is known as constrained clustering and is the main subject of the study presented in this work \cite{triguero2015self}.

\section{Objectives}

The main objective of this work is to provide an overview of the concepts and particularities related to constrained clustering. In order for the new approaches to constrained clustering presented in this work to be properly understood, some of these concepts must be addressed in sufficient depth. It is also in our best interest to provide an example of proper experimental methodology to obtain relevant and subjectivity-free results and conclusions.

@@

Regarding the organization of this work, background knowledge concerning both classic and constrained clustering are presented in Chapters \ref{ch:IntroClustering} and \ref{ch:ConstrainedClustering} respectively. In Chapter \ref{ch:NewProposals} two new approaches to the constrained clustering problem are discussed. The experimental setup and the experimental results of the aforementioned techniques are presented in Chapters \ref{ch:ExperimentalSetup} and \ref{ch:ExperimentalResults} respectively. Afterward, an empirical analysis of these results is performed in Chapter \ref{ch:AnalysisResults}. Finally, conclusions and future work are discussed in Chapter \ref{ch:ConclFW}.
